---
title: "INSAID Task"
author: "unknownda"
date: "2022-10-07"
output: pdf_document
---

## Setting up the data frame 

Installing and loading *readr*, *dplyr*, *readxl*, *caTools*, *ggplot2*, *car*, *caret*, *pROC* and *fastDummies* packages. Then import the weather data from the *Fraud.csv* file.

The data for the task is available in CSV format having 6362620 rows and 10 columns.

The objective is to develop a model for predicting fraudulent transactions for a financial company and use insights from the model to develop an actionable plan.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Loading packages

library(readr)
library(dplyr)
library(caTools)
library(car)
library(ggplot2)
library(caret)
library(pROC)
library(fastDummies)

# Reading the Input file

df <- read_csv("Fraud.csv")
```

## Analyzing Input Data

```{r echo=FALSE}

str(df)
summary(df)

```

The data set contains no missing values. Since no data is present for merchants, hence filtering them out.

```{r echo=FALSE}

df <- df[substr(df$nameDest,1,1)!='M',]

```

* We realize that the old and new balances are not meaningful for the task at hand, rather the change in balance is more important.
* This will also remove possible multi-collinearity.
* Thus creating two new fields that reflect the change in balance in Original and Destination accounts.

```{r echo=FALSE}

df <- df %>% mutate(origChange=oldbalanceOrg-newbalanceOrig, destChange= oldbalanceDest-newbalanceDest)

```

The **Box Plots** are displayed next.

```{r echo=FALSE}

ggplot(df, aes(x=as.factor(isFraud), y=step)) +
  geom_boxplot() +
  labs(title = "Distribution of step by isFraud ")


ggplot(df, aes(x=as.factor(isFraud), y=amount)) +
  geom_boxplot() +
  labs(title = "Distribution of amount by isFraud")

ggplot(df, aes(x=as.factor(isFraud), y=origChange)) +
  geom_boxplot() +
  labs(title = "Distribution of origChange(Change in Origin balance) by isFraud")

```

* No fraud for origChange < 0.
* Checking the same.

```{r echo=FALSE}

F<- df[df$isFraud==1,]
summary(F$origChange)

```

Minimum value of origChange is 0, hence True.

```{r echo=FALSE}

ggplot(df, aes(x=as.factor(isFraud), y=destChange)) +
  geom_boxplot() +
  labs(title = "Distribution of destChange(Change in Destinatiion balance) \nby isFraud")

```

* No fraud for destChange > 0.
* Checking the same.

```{r echo=FALSE}

summary(F$destChange)

```

Maximum value is 315226, considering this point as Outlier.

Filtering out origChange < 0 & destChange > 0.

```{r echo=FALSE}

df <- df[df$origChange>0,]

df <- df[df$destChange<0,]

summary(df)

table(df$isFraud, df$isFlaggedFraud)

```

No isFlaggedFraud = 1, hence this variable is insignificant.

```{r echo=FALSE}

df$type <- as.factor(df$type)

```

## Dummy Variable creation

```{r echo=FALSE}

df_dummy <- dummy_cols(df, select_columns='type') 
summary(df_dummy)
str(df_dummy)

```

* Removing type as it is redundant.
* Removing nameOrig & nameDest.
* Removing old & new balances as their change has been recorded.
* Removing type_TRANSFER AS type_TRANSFER=1 means type_CASH_OUT=type_DEBIT=0

```{r echo=FALSE}

df_dummy <- df_dummy[,-c(2,4:9)]

df_dummy <- df_dummy[,-9]
str(df_dummy)
summary(df_dummy)

```

## Training & Test Set

```{r echo=TRUE, results='hide'}

# Splitting into Training & Test data

set.seed(0)
split <- sample.split(df$nameOrig, SplitRatio = 0.8)
training_set <- subset(df_dummy, split==TRUE)
test_set <- subset(df_dummy, split==FALSE)

```

## Logistic Regression model

Including all variables except isFlaggedFraud.

```{r echo=FALSE}

lr_model <- glm(isFraud~.-isFlaggedFraud, data=training_set, family = binomial)
summary(lr_model)

lr_model_prob <- predict(lr_model, test_set, type="response")

lr_model_pred <- rep(0, 293905)

lr_model_pred[lr_model_prob>0.001] <- 1 

```

Threshold probability is taken as 0.001 so as to achieve maximum Specificity.

```{r echo=FALSE}

confusionMatrix(as.factor(lr_model_pred), as.factor(test_set$isFraud))

roc_1 <- roc(test_set$isFraud, lr_model_pred)
test1 <- roc_1$sensitivities
test2 <- roc_1$specificities

vif(lr_model)

```

* VIF of amount & origChange is very high, hence dropping amount.
* Dropping type_DEBIT as it is not significant.

```{r echo=FALSE}

lr_model2 <- glm(isFraud~step+origChange+destChange+type_CASH_OUT , data=training_set, family = binomial)
summary(lr_model2)

lr_model2_prob <- predict(lr_model2, test_set, type="response")

lr_model2_pred <- rep(0, 293905)

lr_model2_pred[lr_model2_prob>0.001] <- 1 

confusionMatrix(as.factor(lr_model2_pred), as.factor(test_set$isFraud))

roc_2 <- roc(test_set$isFraud, lr_model2_pred)
test3 <- roc_2$sensitivities
test4 <- roc_2$specificities

vif(lr_model2)

```

All VIF within acceptable range but original model is better Predictor.

```{r echo=FALSE}

ggplot() + 
  geom_line(aes(x=test2, y=test1), color="blue") + 
  geom_line(aes(x=test4, y=test3), color="red") +
  labs(title = "ROC curve for the two models") +
  xlab("Specificity") +
  ylab("Sensitivity")

```

Hence Model 1 is better.

## Insights

The key fields that predict fraudulent customer are:
* step 
* amount
* oldbalanceOrg - newbalanceOrig
* oldbalanceDest - newbalanceDest
* Weather the type is CASH OUT or not.

The step field does not make sense as fraudulent transactions can happen at any time and step is a measure of time. All other key fields make sense because these are really the parameters that should indicate a fraud.

## Recommendations

Changes that can be made in the infrastructure:
* The company should record the time based change in the balances rather than actual balances. 
* The company's data collection policy should automatically exclude records where the Origin Balance is increasing wrt time or the Destination Balance is decreasing wrt time.

These changes should save time, effort & memory and this will be reflected by reduced cost.

## Brain-Wave

Let us explore the amount, origChange & destChange fields of all fraudulent records.

```{r echo=FALSE}

str(F)

```

We realize that a lot of fraudulent transactions have the same amount and origChange, so we act accordingly.

```{r echo=FALSE}

df_bw <- df_dummy
logical <- (df_bw$amount - df_bw$origChange)==0
df_bw$isFlaggedFraud[logical]=1
summary(df_bw)

training_set_bw <- subset(df_bw, split==TRUE)
test_set_bw <- subset(df_bw, split==FALSE)

lr_model_bw <- glm(isFraud~., data=training_set_bw, family = binomial)
summary(lr_model_bw)

lr_model_bw_prob <- predict(lr_model_bw, test_set_bw, type="response")

lr_model_bw_pred <- rep(0, 293905)

lr_model_bw_pred[lr_model_bw_prob>0.001] <- 1 

confusionMatrix(as.factor(lr_model_bw_pred), as.factor(test_set_bw$isFraud))

vif(lr_model_bw)

```

* Dropping amount as it has high VIF & is also insignificant.
* Dropping type_DEBIT as it is insignificant.

```{r echo=FALSE}

lr_model_bw2 <- glm(isFraud~. - amount - type_DEBIT, data=training_set_bw, family = binomial)
summary(lr_model_bw2)

lr_model_bw2_prob <- predict(lr_model_bw2, test_set_bw, type="response")

lr_model_bw2_pred <- rep(0, 293905)

lr_model_bw2_pred[lr_model_bw2_prob>0.001] <- 1 

confusionMatrix(as.factor(lr_model_bw2_pred), as.factor(test_set_bw$isFraud))

vif(lr_model_bw2)

```

* All VIF are within control
* So an important change in the company's infrastructure can be the criteria for isFlaggedFraud
* We can set isFlaggedFraud to 1 if (amount - (oldbalanceOrg - newbalanceOrig)) = 0
* This makes the isFlaggedFraud field significant and also increases the model accuracy and specificity.

This is the best model.
